### 2024-07-04 README.Rmd

# MethEvolSIM

<!-- badges: start -->
<!-- badges: end -->

Software for simulation of DNA methylation dynamics on different genomic structures under the basic model for the evolution of DNA methylation along genealogies. The methylation state of each CpG dinucleotide is represented as 1 for unmethylated, 2 for partially methylated and 3 for methylated.

# Installation

You can use the development version of MethEvolSIM from git. The development is being implemented with the `devtools` package. It can be loaded and checked using the following:

```{r}
library(devtools)
load_all()
```

`check()` runs the unit tests. The unit tests that test the functions used for writing output files:
- *validationResults.txt*: Is generated in case an error relative to the characteristics of the objects happens and contains both the error and the object.

# Manual of Use

The initial spatial distribution of CpG sites is given by the user. A data frame with 3 columns named "start", "end" and "globalState" encodes the structural information. Each row provides the information of a different structure. The number of CpG positions is given by $end - start +1$, and each row start value needs to be as the previous row end value + 1. The globalState encodes island and non-island structures with the character value "U" for islands and "M" for non-islands.

```{r}
# Example 3 structures of length 100: non-island, island, non-island
infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"))
infoStr
```

```{r}
# Example 2 islands: first of length 1, second of lenght 15
infoStr <- data.frame(start = c(1, 2),
                        end = c(1, 16),
                        globalState = c("U", "U"))
infoStr
```

Can be sampled as specified in the section "Model parameters" when infoStr has only start, end and globalState columns,
or given by the user as additional columns of the data frame. Note that when given, each of the structures frequencies needs to sum up to 1.

```{r}
# Example 3 structures of length 100 with customized initial methylation state equilibrium frequencies
infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"),
                        u_eqFreq = c(0.1, 0.8, 0.1),
                        p_eqFreq = c(0.1, 0.1, 0.1),
                        m_eqFreq = c(0.8, 0.1, 0.8))
infoStr
```

The default parameter values used by MethEvolSIM can be obtained using the function `get_model_parameters()`, that returns a data frame with the columns named as specified in the section "Model parameters":

The section "Model parameters" provides further information about each of the parameters, including each parameter range of possible values. To simulate data with a different parameterization, the new value of a chosen parameter can be modified in the data frame.

## Initial DNA Methylation Data

### Simulation

The initial data is generated in an object of class combiStructureGenerator.

```{r}
# Example with customized initial methylation frequencies and customized parameter values
custom_infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"),
                        u_eqFreq = c(0.1, 0.8, 0.1),
                        p_eqFreq = c(0.1, 0.1, 0.1),
                        m_eqFreq = c(0.8, 0.1, 0.8))
custom_params <- get_parameterValues()
custom_params$mu <- 0.005
initialD <- simulate_initialData(infoStr = custom_infoStr, params = custom_params)
# Returns customized parameters and simulated data
initialD$params
initialD$data
```

```{r}
# Example with simulated initial methylation frequencies and default parameter values
custom_infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"))
initialD <- simulate_initialData(infoStr = custom_infoStr)
# Returns default parameters
initialD$params
initialD$data
```

To get the parameter values of a combiStructureGenerator object, it can be given as argument to the function `get_parameterValues()`

```{r}
combiStr_object <- initialD$data
get_parameterValues(combiStr_object)
```

### Exploration of initial data

The methods of the class combiStructureGenerator can be used to access each of the contained structures:

```{r}
# E.g. access fist contained structure
combiStr_object$get_singleStr(1)
```

Each contained structure is an object of singleStructureGenerator and its methods can be used to get the contained information.

```{r}
# E.g. get the equilibrium frequencies of the first singleStructureGenerator
combiStr_object$get_singleStr(1)$get_eqFreqs()
```

```{r}
# E.g. get the sequence of methylation states of the second singleStructureGenerator
combiStr_object$get_singleStr(1)$get_seq()
```

## Simulate DNA Methylation Data Evolution

The evolution data is generated in an object of class treeMultiRegionSimulator.

```{r}
# Example with customized initial methylation frequencies, customized parameter values and default dt (0.01)
tree <- "((a:1, b:1):2, c:2, (d:3.7, (e:4, f:1):3):5);"
custom_infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"),
                        u_eqFreq = c(0.1, 0.8, 0.1),
                        p_eqFreq = c(0.1, 0.1, 0.1),
                        m_eqFreq = c(0.8, 0.1, 0.8))
custom_params <- get_parameterValues()
custom_params$mu <- 0.005
evolD <- simulate_evolData(infoStr = custom_infoStr, tree = tree, params = custom_params)
# Returns customized parameters, tree used, time step length for SSE process used (dt) and simulated data
evolD$data
evolD$dt
evolD$tree
evolD$params
```

If the evolution of given data at the root is simulated, and it wants to be done with customized parameter values, then the customized parameters are given to `simulate_initialData()` instead to `simulate_evolData()`.

```{r}
# Example with customized initial methylation frequencies, customized parameter values and default dt (0.01)
tree <- "((a:1, b:1):2, c:2, (d:3.7, (e:4, f:1):3):5);"
custom_infoStr <- data.frame(start = c(1, 101, 201),
                        end = c(100, 200, 300),
                        globalState = c("M", "U", "M"),
                        u_eqFreq = c(0.1, 0.8, 0.1),
                        p_eqFreq = c(0.1, 0.1, 0.1),
                        m_eqFreq = c(0.8, 0.1, 0.8))
custom_params <- get_parameterValues()
custom_params$mu <- 0.005
initialD <- simulate_initialData(infoStr = custom_infoStr, params = custom_params)$data
evolD <- simulate_evolData(rootData =initialD, tree = tree)
```
### Exploration of evolved data

The treeMultiStructureGenerator object contains a list `$Branch` with the combiStructureGenerator representing the data at the each node after a tree branch (but for the data at the root).

To know which offspring branches a node has, the combiStructureGenerator object method `$get_offspring_index()` can be used.

```{r}
treeData <- evolD$data
# E.g. offspring of data at root 
treeData$Branch[[1]]$get_offspring_index()
```
If the node is an inner node, its name is set as NULL. It can be accessed with the combiStructureGenerator method `$get_name()`

```{r}
treeData$Branch[[1]]$get_name()
```
But the leafs are named as in the given tree.

```{r}
treeData$Branch[[4]]$get_name()
```
If there were IWE events sampled in a branch, it can be accessed with the combiStructureGenerator method `$get_IWE_events()`

```{r}
treeData$Branch[[4]]$get_IWE_events()
```
In the case we want to know the equilibrium frequencies of a singleStructureGenerator object contained in the combiStructureGenerator we can use the combiStructureGenerator method to access the singleStructureGenerator object of interest `get_singleStr(i)` followed by the singleStructureGenerator method `get_eqFreqs()`

```{r}
# E.g. get the ewuilibrium frequencies of the second structure at the leaf "b"
treeData$Branch[[4]]$get_singleStr(2)$get_eqFreqs()
```

@@

Structures are categorized as CpG islands or non-islands. In the root of the genealogy we assume that each structure samples an equilibrium probability triple $(\pi_u,\pi_p, \pi_m)$. Current parameter values favor prevalent homogeneity in methylation levels within structures, most CpG islands with a majority predominantly unmethylated and a small proportion predominantly methylated, and most non-islands with a majority predominantly methylated and a small proportion predominantly unmethylated. The sampling of the equilibrium probability tripple is done using two distinct Beta distributions to sample the corresponding methylation frequencies.

@@

Following the same rationale as for island parameterization, all the regions not belonging to an island sample a common equilibrium probability triple from two Beta distributions with parameters $\alpha_{NI_p}$, $\beta_{NI_p}$, $\alpha_{NI_m}$ and $\beta_{NI_m}$. The parameter values are currently set to favor the characteristic high overall methylation level.

You can simulate DNA methylation along a tree given some data at root, or by simulating data at root given a structural distribution and letting it evolve.

**Parameter ranges** 
The alpha and beta parameters of a Beta distribution must be greater than 0, that is (0, $\infty$).

@@

**IWE parameterization: mu. Parameter range:**
Should be a small number, but it can take any value from 0, that is [0, $\infty$)

@@

As SSEc transition probabilities are dependent on adjacent sites we need to approximate SSE based evolution by short time steps, allowing for CpG sites to actualize their state. Therefore, we first discretize the branch intervals between IWEs in time intervals of length 0.01. For each CpG position and each time interval the rate matrix for the transitions between the states unmethylated, partially methylated and methylated results from the addition of the rate matrices of the two types of SSE events, so that $\mathbb{E}(R_i + R_c)=1$, $R_i$ and $R_c$ representing the rate factors of independent and correlated processes. The proportion of each of the two SSE types is determined by the parameter $\iota$.

@@

**SSE process parameterization: iota, alpha_Ri**
**Parameter ranges** 
The alpha parameter of a Gamma distribution must be greater than 0, that is (0, $\infty$). Iota must be between 0 (non-included) and 1, that is (0,1].

>>>>

```{r setup}
library(MethEvolSIM)
```

MethEvolSIM is an R package for simulating DNA methylation dynamics
on different genomic structures along a given tree.
It allows simulating methylation and demethylation events at loci
within two different types of genomic structures: regions with low methylation or with high methylation.
The data can be interpreted as:

1- CpG sites inside or outside CpG islands or

2- any type of methylation site in regions of higher or lower methylation or

3- CpG islands inside or outside genomic archipelagos (regions with many CpG islands).

In this document, unless specified otherwise, we use the first interpretation.

# A first example
As a simple example with default parameter values we show a simulation of
evolution of a simplified genomic structure on a tree:

```{r}
# Define the tree:
tree <- "((a:1, b:1):2, c:2);"
# Define the genomic structure. Here 1 island ("U") containing 100 CpGs 
# surrounded by 2 non-island structures ("M") each containing 10 CpGs
infoStr <- data.frame(n = c(10, 100, 10),
                      globalState = c("M", "U", "M"))
# Simulate 1 replicate of data at the tree tips
output <- simulate_evolData(infoStr = infoStr, tree = tree)
```

In the simulation output 0 stands for unmethylated, 1 for methylated and 0.5 for
partially methylated or hemimethylated.

```{r}
# In replicate 1, tip 2: name of the tip of the tree
output$data[[1]][[2]]$name
# In replicate 1, tip 2: sequence of methylation states of left non-island structure
output$data[[1]][[2]]$seq[[1]]
# In replicate 1, tip 2: sequence of methylation states of island structure
output$data[[1]][[2]]$seq[[2]]
# In replicate 1, tip 2: sequence of methylation states of right non-island structure
output$data[[1]][[2]]$seq[[3]]
```

@@

Structures are categorized as CpG islands or non-islands (alternatively regions with low/high methylation or archipelagos/non-archipelagos). In the root of the genealogy we
assume that each structure samples an equilibrium probability triple $(\pi_u,\pi_p, \pi_m)$ for the methylation states: unmethylated, partially-methylated and methylated, respectively.
Current parameter values favor prevalent homogeneity in methylation levels within structures,
most CpG islands predominantly unmethylated and a small proportion predominantly
methylated, and most non-islands predominantly methylated and a small
proportion predominantly unmethylated.
The sampling of the equilibrium probability triple is done using two distinct Beta
distributions to sample the corresponding methylation frequencies.

@@

Following the same rationale as for island parameterization, each non-island structure 
samples a common equilibrium probability triple from two Beta distributions
with parameters $\alpha_{NI_p}$, $\beta_{NI_p}$, $\alpha_{NI_m}$ and $\beta_{NI_m}$.
The parameter values are currently set to favor the characteristic high overall methylation level.

**Parameter ranges** 
The alpha and beta parameters of a Beta distribution must be greater than 0,
that is in (0, $\infty$).

@@

**IWE parameterization: mu. Parameter range:**
Should be a small number, but it can take any value from 0, that is in $[0, \infty)$

@@

As SSEc transition probabilities are dependent on adjacent sites we use Gillespie's $\tau$-leap approximation.
to allow for CpG sites to o update their neighbouring state and rates of change after short time steps.
Therefore, we first discretize the branch intervals between IWEs in time intervals
of default length $dt =0.01$.
For each CpG position and each time interval the rate matrix for
the transitions between the states unmethylated, partially methylated and methylated
results from the addition of the rate matrices of the two types of SSE events,
so that $\mathbb{E}(R_i + R_c)=1$, $R_i$ and $R_c$ representing the rate factors of
independent and correlated processes. The independent process samples the new methylation state from the structure's equilibrium frequencies and the correlated process copies the left or right neighbor state with equal probabilities.
The proportion of each of the two SSE types
is determined by the model parameter $\iota$.

@@

**SSE process parameterization: iota, alpha_Ri**
**Parameter ranges** 
The alpha parameter of a Gamma distribution must be greater than 0, that is (0, $\infty$).
Iota must be between 0 (non-included) and 1, that is in (0,1].

# Simulation Setup

The initial spatial distribution of CpG sites in islands and non-island structures is given by the user through a data frame with 2 columns named "n" and "globalState"
, which encode the structural information. 
Each row provides the information of a different genomic structure.

- **n** number of positions for each structure (e.g. number of CpG sites).

- **globalState** encodes island and non-island structures with the character
value "U" for unmethylated regions (e.g. CpG islands) and "M" for methylated regions (e.g. non-CpG islands).

```{r}
# Example 3 structures of length 100 each:
# one non-island containing 100 CpG sites,
# one island containing 100 CpG sites,
# and one non-island containing 100 CpG sites
infoStr <- data.frame(n = c(100, 100, 100),
                        globalState = c("M", "U", "M"))
infoStr
```

```{r}
# Example for 2 islands: first of length 1, second of length 15
infoStr <- data.frame(n = c(1, 15),
                      globalState = c("U", "U"))
infoStr
```

The initial methylation state distribution assigns a triple of methylation equilibrium frequencies() $\pi_u$ for unmethylated, $\pi_p$ for partially-methylated, and $\pi_m$ for methylated) to each of the simulated genomic structures. This assignation can be done in two ways:

1 - Sampled as specified in the section "Model parameters: Initial methylation state distribution" when the given `infoStr` only contains `n` and `globalState` columns.

2 - Determined by the user providing in `infoStr` additional columns: `u_eqFreq` for $\pi_u$, `p_eqFreq` for $\pi_p$ and `m_eqFreq` for $\pi_m$. Note that when given, the triple of frequency values for each of the structures needs to sum up to 1, so that for each row $i$, $\pi_u(i)+\pi_p(i)+\pi_m(i)=1$.

```{r}
# Example 3 structures of length 100 with customized initial methylation state
# equilibrium frequencies
infoStr <- data.frame(n = c(100, 100, 100),
                      globalState = c("M", "U", "M"),
                      u_eqFreq = c(0.1, 0.8, 0.1),
                      p_eqFreq = c(0.1, 0.1, 0.1),
                      m_eqFreq = c(0.8, 0.1, 0.8))
infoStr
```

The default parameter values used by MethEvolSIM can be obtained using the
function `get_parameterValues()`, which returns a data frame with the columns named
as specified in the section "Model parameters":

Additionally, it returns the three rates of change for SSEi process,`Ri1`, `Ri2`, and `Ri3` (see section "Model Parameters: Single-site events (SSEs)").

The section "Model parameters" provides further information about each of the parameters,
including the range of possible values for each parameter.
To simulate data with a different parameterization, the new value of the chosen parameter(s) can be modified in the data frame, but for the cases of `Ri1`, `Ri2`, and `Ri3`, which are modified through `iota` and `alpha_Ri`.

Additional details regarding the `get_parameterValues()` function can be found in its documentation.

# Simulation of DNA Methylation Data

DNA Methylation Data is simulated using three R6 classes: `singleStructureGenerator`, `combiStructureGenerator`, and `treeMultiRegionSimulator`. 
Instances of the `singleStructureGenerator` class represent individual genomic structures (e.g. islands or non-islands) and are encapsulated within instances of the `combiStructureGenerator` class, which, in turn, represent the simulated data structure (e.g. a sequence consisting of a non-island containing 100 CpGs, an island containing 100 CpGs, and a non-island containing 100 CpGs). Finally, instances of the `treeMultiRegionSimulator` class contain a `combiStructureGenerator` object at each node of a given tree, including its root and leaves.

Generally, simulations are conducted as explained in subsection "Simulation of Data Evolution", with a documented user-interface function explaining the output structure: `simulate_evolData()`. Alternativelly, the user can simulate initial data so that the root of the tree is fixed as explained in subsection "Simulation of Initial Data". The latter option generates objects of class `combiStructureGenerator` containing objects of class `singleStructureGenerator`. In the subsection "Simulation of Initial Data: Exploration of Initial Data" an example to interact with the classes is given, while further information about the attributes and methods of each class is provided in their documentation.

## Simulation of Initial Data

The initial data can be generated as an object of class `combiStructureGenerator`.

```{r}
# Example with customized initial methylation frequencies and customized parameter values
custom_infoStr <- data.frame(n = c(100, 100, 100),
                             globalState = c("M", "U", "M"),
                             u_eqFreq = c(0.1, 0.8, 0.1),
                             p_eqFreq = c(0.1, 0.1, 0.1),
                             m_eqFreq = c(0.8, 0.1, 0.8))
custom_params <- get_parameterValues()
custom_params$mu <- 0.005
initial_customD <- simulate_initialData(infoStr = custom_infoStr, params = custom_params)
# Returns customized parameters and simulated data
initial_customD$params
initial_customD$data
```

```{r}
# Example with sampled initial methylation frequencies and default parameter values
custom_infoStr <- data.frame(n = c(100, 100, 100),
                             globalState = c("M", "U", "M"))
initialD <- simulate_initialData(infoStr = custom_infoStr)
# Returns default parameters
initialD$params
initialD$data
```

To get the parameter values of a `combiStructureGenerator` object, it can alternativelly be given as argument to the function `get_parameterValues()`.

```{r}
combiStr_object <- initialD$data
get_parameterValues(rootData = combiStr_object)
```

### Exploration of Initial Data

The methods of the class `combiStructureGenerator` can be used to access each of the contained structures:

```{r, eval=FALSE}
# E.g. access fist structure (non-island)
combiStr_object$get_singleStr(1)
```

Each contained structure is an object of `singleStructureGenerator` and its methods can be used to get the contained information.

```{r}
# E.g. get the methylation equilibrium frequencies of the first singleStructureGenerator
combiStr_object$get_singleStr(1)$get_eqFreqs()
```

```{r}
# E.g. get the sequence of methylation states of the second singleStructureGenerator
# Encoded as: 1 for unmethylated, 2 for partially-methylated, 3 for methylated
combiStr_object$get_singleStr(2)$get_seq()
```

## Simulation of Data Evolution

The evolution data is generated with the function `simulate_evolData()`. It returns  the parameters used (`$params`), the length of the time step used for the Gillespie's $\tau$-leap approximation (`$dt`, default 0.01), the tree used (`$tree`), and the simulated data (`$data`). The description of the simulated data is available in the function's documentation.

It takes as parameters:

- `tree`: A string representing a tree in Newick format.

- `infoStr`: A dataframe with the information of the number of CpGs corresponding to each island and non-island region (columns named `n` and `globalState`). Additionally the initial methylation frequencies can be specified adding the columns `u_eqFreq`, `p_eqFreq` and `m_eqFreq`.

- `dt`: The length of time steps for Gillespie's $\tau$-leap approximation. Numerical, default is 0.01.

- `n_rep`: The number of replicates. Integer, default is 1.

- `only_tip`: Logical indicating whether to extract data only for tips (TRUE, default) or to extract the information for all the tree branches (FALSE).

- `params`: Default NULL for simulation with default parameter values. When customized parameter values are used the output dataframe of `get_parameterValues()` can be modified and given to the function. E.g:

```{r}
# Example with customized initial methylation frequencies, customized parameter values 
# and default dt (0.01)
tree <- "((a:1, b:1):2, c:2, (d:3.7, (e:4, f:1):3):5);"
custom_infoStr <- data.frame(n = c(100, 100, 100),
                             globalState = c("M", "U", "M"),
                             u_eqFreq = c(0.1, 0.8, 0.1),
                             p_eqFreq = c(0.1, 0.1, 0.1),
                             m_eqFreq = c(0.8, 0.1, 0.8))
custom_params <- get_parameterValues()
custom_params$mu <- 0.005
evolD <- simulate_evolData(infoStr = custom_infoStr, tree = tree, params = custom_params, n_rep = 3, only_tip = TRUE)
```

```{r, eval=FALSE}
# Returns customized parameters, tree used, time step length for SSE process used (dt) 
# and simulated data
evolD$data
evolD$dt
evolD$tree
evolD$params
```

```{r}
# The simulated data is returned as a list. Each element of the list corresponds to a
# simulation replicate:
length(evolD$data)
rep1 <- evolD$data[[1]]
rep2 <- evolD$data[[2]]
rep3 <- evolD$data[[3]]
```

```{r}
# When only_tip is TRUE, each replicate contains for each tip: its name and the 
# sequence of methylation states. E.g.
rep1_tip1 <- rep1[[1]]
rep1_tip1$name
# Seq is a list with the methylation states of each simulated structure.
# In this example, 3 structures: non-island, island and non-island ...
length(rep1_tip1$seq)
# Each with 100 CpGs
length(rep1_tip1$seq[[1]])
```

When simulations are conducted with `only_tip = FALSE` the `$data` output contains more information.

```{r}
tree <- "((a:1, b:1):2, c:2, (d:3.7, (e:4, f:1):3):5);"
custom_infoStr <- data.frame(n = c(100, 100, 100),
                             globalState = c("M", "U", "M"),
                             u_eqFreq = c(0.1, 0.8, 0.1),
                             p_eqFreq = c(0.1, 0.1, 0.1),
                             m_eqFreq = c(0.8, 0.1, 0.8))
custom_params <- get_parameterValues()
custom_params$mu <- 0.005
evolD <- simulate_evolData(infoStr = custom_infoStr, tree = tree, params = custom_params, n_rep = 3, only_tip = FALSE)
```

```{r}
# When only_tip is FALSE. $data output contains the simulated data and the 
# information on the relationship between tree branches.
names(evolD$data)
```

```{r}
# The information of the tree branches can be accessed using $branchInTree. 
treeStr <- evolD$data$branchInTree
# It is a list in which each element index represents the index of the branch
root <- treeStr[[1]] # First branch (or tree root)
b2 <- treeStr[[2]] # Second branch 
# ...
# Each branch contains information of the offspring and parent indexes.
root$parent_index # The tree root does not have parent branches
root$offspring_index # The tree root has 3 daughter branches: 2, 5 and 6. 
b2$parent_index # Branch 2 therefore has as parent branch the root.
b2$offspring_index # Branch 2 also has 2 daughter branches: 3 and 4
```

```{r}
# The simulated data can be accessed using $sim_data.
sim_data <- evolD$data$sim_data
# As before, each list element corresponds to a simulation replicate. 
# E.g. replicate 3
rep3 <- sim_data[[3]]
# In each replicate, each element of the list corresponds to a tree branch
# the indexes correspond to the information in $branchInTree
root <- rep3[[1]]
b2 <- rep3[[2]]
b3 <- rep3[[3]]
```

```{r}
# Each tree branch contains:
# - the branch name (NULL for tree root and inner nodes and 
# tip name for the tree tips):
root$name
b2$name
b3$name
```
```{r}
# - the information on IWE events that happened in that branch. 
root$IWE # The root always has NULL because its branch length is 0
# The rest of the branches have FALSE when no IWE happened or a list containing
# $islands corresponds to the island index(es) to which the event(s) happened 
# $times corresponds to the branch time in which the event happened.
b2$IWE
b3$IWE
```

```{r}
# - A list with each element representing the sequence of methylation states of 
# each simulated island and non-island
# Encoded as 0 for unmethylated, 0.5 for partially methylated and 1 for methylated
root$seq[[1]] # First sequence (non-island)
root$seq[[2]] # Second sequence (island)
```

```{r}
# - A list with each element representing the methylation frequencies for the states: unmethylated, partially-methylated and methylated
root$eqFreqs[[1]][1] # First structure of root (non-island) equilibrium frequency for unmethylated state
root$eqFreqs[[1]][2] # First structure of root (non-island) equilibrium frequency for partially-methylated state
root$eqFreqs[[1]][3] # First structure of root (non-island) equilibrium frequency for methylated state
```

- `rootData`: Alternative argument to `infoStr` (only one can be given). It enables simulating replicates of evolution along a tree with fixed data at root. When used, it takes the output of `simulate_initialData()$data`. If customized parameter values are to be used, then the customized parameters are given to `simulate_initialData()` instead to `simulate_evolData()`.

```{r}
# Example with customized initial methylation frequencies, customized parameter values and default dt (0.01)
tree <- "((a:1, b:1):2, c:2, (d:3.7, (e:4, f:1):3):5);"
custom_infoStr <- data.frame(n = c(100, 100, 100),
                             globalState = c("M", "U", "M"),
                             u_eqFreq = c(0.1, 0.8, 0.1),
                             p_eqFreq = c(0.1, 0.1, 0.1),
                             m_eqFreq = c(0.8, 0.1, 0.8))
custom_params <- get_parameterValues()
custom_params$mu <- 0.005
initialD <- simulate_initialData(infoStr = custom_infoStr, params = custom_params)$data
evolD <- simulate_evolData(rootData =initialD, tree = tree)
```
